---
title: "NYC Airbnb Price Predictions"
author: "Alex Shen, Daniel Zhou"
output:
  pdf_document: default
  html_document: default
code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r loading_packages, include = FALSE, warning = FALSE}
library(tidyverse)
library(broom)
library(patchwork)
```


## Reading and cleaning data

Our data set was obtained from Kaggle, and contains information about Airbnb 
listings for NYC in 2019. We want to see what factors can best predict the 
price of an Airbnb listing in NYC by creating a linear regression model with 
price as the response variable. The variables that could be used to create a 
linear regression at first glance are price (presumably per day/night in 
dollars), neighborhood_group (borough), neighborhood, latitude, longitude, 
room_type (shared, private, or entire house/apartment), minimum_nights (minimum 
nights required), number_of_reviews, last_review (date of last review), 
reviews_per_month, calculated_host_listings_count (number of listings from the 
same host), and availability_365 (number of days available).


```{r data, warning = FALSE}
AB <- read.csv("data/AB_NYC_2019.csv")
AB <- AB %>% 
  filter(price>0) %>% 
  mutate(neighborhood_group = as.factor(neighbourhood_group),
         neighbourhood = as.factor(neighbourhood),
         room_type = as.factor(room_type)) %>% 
  select(neighborhood_group, neighbourhood, latitude, longitude, room_type, 
         price, minimum_nights, number_of_reviews, last_review, 
         reviews_per_month, calculated_host_listings_count, availability_365)
AB <- na.omit(AB)
```

After reading in the data, we do some cleaning necessary to conduct our linear 
regression. We select the variables that would be useful in conducting 
regression, make the necessary variables factor variables, and filter for 
prices greater than 0, since having prices equal to 0 does not make much sense,
and would break the code later on. 


```{r price_dist, warning = FALSE}
ggplot(data = AB, mapping = aes(x = price)) + 
  geom_histogram(binwidth = 20, fill = "steelblue") +  xlim(0, 2500) +
  labs(title = "Distribution of Price", x = "Price ($)", y = "Count")
```

Here we can see the distribution of price. As we can see, the vast majority of 
prices are between 0 and 300 dollars, and the data is skewed right with what 
looks like many outliers visually on the right tail. This makes sense for price 
to be concentrated at such prices (what average people can afford) and have a 
few outliers of super luxurious Airbnb's.



```{r cleaning, warning = FALSE}
AB_new <- AB %>% 
  filter(price > quantile(AB$price,.25) - 1.5*IQR(AB$price) & 
           price < 1.5*IQR(AB$price) + quantile(AB$price,.75)) %>% 
  mutate(log_price = log(price))

ggplot(data = AB_new, mapping = aes(x = price)) + 
  geom_histogram(binwidth = 20, fill = "steelblue") +  xlim(0, 2500) + 
  labs(title = "Distribution of Price Without Outliers", x = "Price ($)", 
       y = "Count")

```

As a result, we eliminate price outliers from the data set. We also log 
transformed price to prevent negative price predictions from our model (and to 
help out the constant variance condition, discussed later). If we did not 
filter for price is greater than 0, this would not be possible. 


## Creating our Model

First, we wanted to think about what interaction effects we should include in 
our full model before being backwards selected. We thought that of all the 
variables, there could be an interaction between minimum nights and room type, 
so we decided to investigate.

```{r interactions, message = FALSE, warning = FALSE}
ggplot(data = AB_new, aes(x = minimum_nights, 
                          y = log_price, color = room_type)) + geom_point() + 
  geom_smooth(method = lm) + 
  labs(title = "Minimum nights and log price relationship", 
       subtitle = "Separated by room type", x = "Minimum Nights", 
       y = "Log Price") + scale_color_manual(values = c("royalblue2", 
                                                        "steelblue", 
                                                        "darkblue"))
```

Looking at the graph, we can tell that based on the level of room type, the 
relationship (slope) between minimum nights and log price changes. As a 
result there is an interaction between room type and minimum nights, so we 
will add this interaction effect to the full model.


```{r full_model, warning = FALSE}
int_only <- lm(log_price ~ 1, data = AB_new)
full_model <- lm(log_price ~ neighborhood_group + minimum_nights + room_type + 
                   number_of_reviews + reviews_per_month + 
                   calculated_host_listings_count + availability_365 + 
                   room_type*minimum_nights, data = AB_new)
```


```{r select_model, warning = FALSE}
Selected_model <- step(full_model, scope = formula(int_only), 
                       direction = "backward", trace = 0)
tidy(Selected_model) %>%
  knitr::kable(digits = 5)
save(Selected_model, file="R/selected_model.Rdata")
```

We chose to use a backwards selection by AIC rather than BIC, because AIC 
usually lends itself more towards prediction tasks, which is what we are trying 
to do. We create our full model of all possible variables that we deem useful, 
then remove each variable that minimizes AIC until it is no longer possible. 

After looking at this model, we see that all of the coefficients have really 
low p values except for neighborhood_group Staten Island and number_of_reviews. 
Because neighborhood_group Staten Island is a level of the neighborhood_group, 
it means that the price of Airbnb's in Staten Island is not significantly 
different than those in the base group (Bronx). Because the coefficient is so 
small and because it is a level of the larger neighborhood_group, we will keep 
it in the model. For number_of_reviews, the p-value is pretty much equal to our
chosen significance cutoff of 0.05, so we decide to keep it in the model. 

One thing to note is that our model has a log transformed response variable. 
So, an increase in one of the dependent variables (with a coefficient B) by 1 
means that the predicted median price will be multiplied by a factor of e^B. 

Next, we will check conditions. 

## Conditions

```{r std resid plot, warning = FALSE}
AB_aug <- augment(Selected_model) %>%
   mutate(obs_num = row_number()) 

ggplot(data = AB_aug, mapping = aes(x = .fitted, y = .std.resid)) + 
  geom_point(color = "steelblue") +
  geom_hline(yintercept = 0, color = "red") +
  labs(title = "Standardized Residual Plot for Price Model", 
       x = "Log Predicted Price", y = "Standardized Residual")
```

From our graph, we can see that there is a similar amount of positive and 
negative residuals, which satisfies the linearity condition. The magnitude of 
the residuals does seem to change a bit as price increases. However, this was 
much more pronounced before we did the log transformation. As a result, the 
constant variance condition is not completely satisfied, but we will proceed 
with caution.



```{r normality_check, warning = FALSE}
resid_hist <- ggplot(data = AB_aug, aes(x = .std.resid)) +
  geom_histogram(fill = "steelblue") +
  labs(x = "Standardized Residuals", title = "Dist. of Residuals")
resid_qq <- ggplot(data = AB_aug, aes(sample = .std.resid)) +
  stat_qq(color = "steelblue") + 
  stat_qq_line(color = "red") +
  labs(title = "Normal QQ-plot of Std. Resid.")
resid_hist + resid_qq
```

The normality condition is satisfied because the distribution of the 
standardized residuals is approximately normal, seen in our histogram and in 
our normal QQ plot. 


## Result discussions

Going into the analysis, we did not have too much of an idea of which variables 
would be influential in predicting the prices of Airbnb's, except that we 
thought that neighborhood_group and room_type would matter (which was correct).

After looking at the results, most of the results were unsurprising. The base 
room_type of entire apartment/home usually costs more than a single room, which
usually costs more than a double room. The number of listings a host had 
(calculated_host_listings_count) and the availability of the Airbnb 
(availability_365) both had small but positive coefficients, showing that 
as each of them increased generally the price would increase. We found that 
the the order of boroughs for median predicted price of Airbnb's from most 
expensive to least expensive is Manhattan, Brooklyn, Queens, Staten Island, 
and Bronx, with Staten Island and Bronx being very similar. For the minimum 
required nights, as minimum_nights increased price tended to decrease. This 
makes sense because the more nights you have to pay for, the lower the price
can be (similar to buying things in bulk). Finally, it seemed that the number 
of reviews seemed to have a slight negative relationship with price. This did 
not make much sense, because it would seem like a lot of reviews for a host 
generally means a good Airbnb which can be priced higher. However, a lot of 
reviews for an Airbnb means that many people were able to stay, meaning the 
Airbnb might be in the more affordable range of prices rather than being one 
of the more expensive ones. As a result, after further thought, this result 
kind of makes sense as well. 

We will use this model to create an RShiny app that allows users to enter in 
the information about an Airbnb and our app will predict the  price. 